LHC abort gap monitor electronics upgrade

### 2024-05-11 paper.tex
Update on Overleaf.

\subsection{Operational system, used since 2010}
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{existing_system_processing.pdf}
    \caption{Operational AGM system overview}
    \label{fig:current_system_processing}
\end{figure}

The currently used system is shown in Fig.~\ref{fig:current_system_processing}.
%
The optical path, including filters and MCP-PMT, is common to both the operational version of the system, and the proposed upgrade.
%As well, the PMT type is the same in both systems.
The operational setup utilizes a digital acquisition board (DAB64x) incorporating an Altera Stratix EP1S40 FPGA for data processing.
%
The DAB64x hosts two Individual Bunch Measurement Systems (IBMS)~\cite{ibms}, integrating the analog input signal via the LHCb2002 ASIC~\cite{lhcb_asic}.
%
The analog signal generated by  the MCP-PMT is amplified  prior to integration to  enhance the measurement's Signal-to-Noise Ratio  (SNR).
%
The amplifier (HAMAMATSU C5594~\cite{bsra_first_operation}) has a fixed voltage gain of \SI{36}{dBV} and the bandwidth of \SI{1.5}{GHz} with the low cutoff frequency at \SI{50}{kHz}.
%
The amplified signal is fed to the LHCb2002 in the IBMS integrating in \SI{25}{\nano\second} intervals, producing a stream of analog  'DC-like' values.
%
These are subsequently sampled by a \SI{40}{MSPS} 14-bit ADC, resulting in a digital data stream, where each value corresponds to an integral of an individual LHC \SI{25}{ns} bucket.
%
The DAQ gateware sums four consecutive samples to compute the
proton charge present on a~\SI{100}{ns} interval.
%
This process generates 64 values per single turn acquisition.
%
For each individual value, a standard deviation and mean value is calculated, and separately filtered by EMA. 
%
All collected data are stored in a RAM block shared to a front-end computer (FEC) via VME64 bridge.
%
A dedicated  front-end  server, publishes the data stream through FESA~\cite{fesa}.

@@

Figure \ref{fig:amp_schematics} depicts the block diagram of the new amplifier.
%
The amplifier was designed to match the AGM MCP-PMT and to be  highly configurable.
% A = U^2/(i^2 * R^2) => A_dB = 20*log10(U/(i*R))
%
A universal asynchronous receiver-transmitter (UART) interface allows for setting a~trans-impedance gain.
%
The gain is adjustable in a~range of \SI{5.8}{mV/\micro A} to \SI{1.1}{V/\micro A}.
%
This corresponds to a~voltage gain of \SI{41}{dB} to \SI{87}{dB} in a $50\Omega$ system.
%
A programmable DC-offset compensation is implemented in the amplifier.
%
It can be used - together with variable gain - to optimize the signal's amplitude for the ADC.
%
The amplifier design incorporates an alarm. This alarm interlocks the MCP-PMT gate when the incident light produces a current nearing the MCP-PMT's maximum allowable average current.
%The amplifier implements 
%an alarm that interlocks the MCP-PMT gate when the incident light generates a current approachng  the MCP-PMT's maximum allowable average current.
%
The alarm threshold setting is stored through the UART into the amplifier's non-volatile memory, so that 
%it protects 
the MCP-PMT is protected also after a power reset. %immediately at the amplifier power-up.
%
%The alarm status is also transmitted to the gateware.
%
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{amp_schematics.pdf}
    \caption{New AGM custom amplifier.}
    \label{fig:amp_schematics}
\end{figure}
%
\subsection{Acquisition system}
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_daq.pdf}
    \caption{Acquisition system schematics.}
    \label{fig:bsra_daq}
\end{figure}
%
The new DAQ (Fig.~\ref{fig:bsra_daq}) receives asynchronous data samples as a stream of four frames every~\SI{8}{\nano s}, each corresponding to one channel.
%
Each frame contains 4$\times$16-bit words, storing left-aligned 14-bit ADC samples. 
%
To compute integrals over 4$\times$~\SI{25}{\nano\second} bunch slots, the system employs data tagging. 
%
The tagger resamples the LHC timing, and generates timing masks that  provide each individual ADC frame with %allowing for identification of 
bunch and turn start markers. 
%within each individual ADC frame.
%
Tags are inserted into the  2 unused least significant bits (LSBs) of each word to indicate the start of the  the bunch or turn.
\\
%
The tagged samples are stored in a~FIFO (First In, First Out), and then integrated. 
%
The integrator is triggered %begins summing values 
by the first bunch tag appearing within the acquisition window.
%
In parallel the system reads a stream of 64~values of a stored 'offset' from a pedestal memory (P[63:0]).
%
The offset is subtracted  from the integrated data stream on the fly.
%
Such pedestal subtraction is needed to remove systematic effects, e.g. related to the  MCP-PMT gate opening. 
%
The output values are filtered by EMA, the results are stored  in RAM, and then read out through the VME64x bus~\cite{my_thesis}.

>>>>

\subsection{Operational system, used since 2010}
The currently used system is shown in Fig.~\ref{fig:current_system_processing}.
%
The optical path, including filters and MCP-PMT, is common to both the operational version of the system, and the proposed upgrade.
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{existing_system_processing.pdf}
    \caption{Operational AGM system overview}
    \label{fig:current_system_processing}
\end{figure}
%As well, the PMT type is the same in both systems.
The operational setup utilizes a digital acquisition board (DAB64x) incorporating an Altera Stratix EP1S40 FPGA for data processing.
%
The DAB64x hosts two Individual Bunch Measurement Systems (IBMS)~\cite{ibms}, integrating the analog input signal via the LHCb2002 ASIC~\cite{lhcb_asic}.
%
The analog signal generated by  the MCP-PMT is amplified  prior to integration to  enhance the measurement's Signal-to-Noise Ratio  (SNR).
%
The amplifier (HAMAMATSU C5594~\cite{bsra_first_operation}) has a fixed voltage gain of \SI{36}{dBV} and the bandwidth of \SI{1.5}{GHz} with the low cutoff frequency at \SI{50}{kHz}.
%
The amplified signal is fed to the LHCb2002 in the IBMS integrating in \SI{25}{\nano\second} intervals, producing a stream of analog  'DC-like' values.
%
These are subsequently sampled by a \SI{40}{MSPS} 14-bit ADC, resulting in a digital data stream, where each value corresponds to an integral of an individual LHC \SI{25}{ns} bucket.
%
The DAQ gateware sums four consecutive samples to compute the
proton charge present on a~\SI{100}{ns} interval.
%
This process generates 64 values per single turn acquisition.
%
For each individual value, a standard deviation and mean value is calculated, and separately filtered by EMA. 
%
All collected data are stored in a RAM block shared to a front-end computer (FEC) via VME64 bridge.
%
A dedicated  front-end  server, publishes the data stream through FESA~\cite{fesa}.

@@

Figure \ref{fig:amp_schematics} depicts the block diagram of the new amplifier.
%
The amplifier was designed to match the AGM MCP-PMT and to be  highly configurable.
% A = U^2/(i^2 * R^2) => A_dB = 20*log10(U/(i*R))
%
A universal asynchronous receiver-transmitter (UART) interface allows for setting a~trans-impedance gain.
%
The gain is adjustable in a~range of \SI{5.8}{mV/\micro A} to \SI{1.1}{V/\micro A}.
%
This corresponds to a~voltage gain of \SI{41}{dB} to \SI{87}{dB} in a $50\Omega$ system.
%
A programmable DC-offset compensation is implemented in the amplifier.
%
It can be used - together with variable gain - to optimize the signal's amplitude for the ADC.
%
The amplifier design incorporates an alarm. This alarm interlocks the MCP-PMT gate when the incident light produces a current nearing the MCP-PMT's maximum allowable average current.
%The amplifier implements 
%an alarm that interlocks the MCP-PMT gate when the incident light generates a current approachng  the MCP-PMT's maximum allowable average current.
%
The alarm threshold setting is stored through the UART into the amplifier's non-volatile memory, so that 
%it protects 
the MCP-PMT is protected also after a power reset. %immediately at the amplifier power-up.
%
%The alarm status is also transmitted to the gateware.
%
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{amp_schematics.pdf}
    \caption{New AGM custom amplifier.}
    \label{fig:amp_schematics}
\end{figure}
%
\subsection{Acquisition system}
The new DAQ (Fig.~\ref{fig:bsra_daq}) receives asynchronous data samples as a stream of four frames every~\SI{8}{\nano s}, each corresponding to one channel.
%
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_daq.pdf}
    \caption{Acquisition system schematics.}
    \label{fig:bsra_daq}
\end{figure}
%
Each frame contains 4$\times$16-bit words, storing left-aligned 14-bit ADC samples. 
%
To compute integrals over 4$\times$~\SI{25}{\nano\second} bunch slots, the system employs data tagging. 
%
The tagger resamples the LHC timing, and generates timing masks that  provide each individual ADC frame with %allowing for identification of 
bunch and turn start markers. 
%within each individual ADC frame.
%
Tags are inserted into the  2 unused least significant bits (LSBs) of each word to indicate the start of the  the bunch or turn.
\\
%
The tagged samples are stored in a~FIFO (First In, First Out), and then integrated. 
%
The integrator is triggered %begins summing values 
by the first bunch tag appearing within the acquisition window.
%
In parallel the system reads a stream of 64~values of a stored 'offset' from a pedestal memory (P[63:0]).
%
The offset is subtracted  from the integrated data stream on the fly.
%
Such pedestal subtraction is needed to remove systematic effects, e.g. related to the  MCP-PMT gate opening. 
%
The output values are filtered by EMA, the results are stored  in RAM, and then read out through the VME64x bus~\cite{my_thesis}.
